Solving Collective Spin Hamiltonians with Decoherence

### 2018-08-22 soc_squeezing/cumulant_expansions.tex
more modifications to the commutator derivation...

where the first terms in the sum over $p$ is

@@

  = n \sum_{k=0}^{m-2} D_{\mu,m-1,n-1}^{(k)}
  \p{2k+1-n+2\mu S_z},
the second

@@

  = n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}

@@

\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} provide a complete expansion for
the commutator $\sp{S_\mu^m,S_\nu^n}_- = C_{\mu mn}$, but the result
is -- to put it lightly -- quite a mess: the expression for

@@

Observing that we always have
$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$, we can rearrange the order
of the sums in \eqref{eq:E_mn_p} and relabel $k_p\to\ell$ to get
\begin{align}
  E_{\mu mn}^{(p)}
  = \f{n!}{\p{n-p}!}
  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+F_{\mu np}}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}},
  \label{eq:E_mn_p_sum}
\end{align}
where for shorthand we define
\begin{align}
  F_{\mu np} \equiv p - n + 2\mu S_z,
  &&
  \sum_{\p{\v k,p,\ell}} X \equiv
  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
  \cdots \sum_{k_{p-1}=k_{p-2}}^\ell X.
\end{align}
We now define
\begin{align}
  f_{\mu np\ell}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{\mu np}},
\end{align}
and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
finding
\begin{align}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  \prod_{q=1}^r f_{\mu np\ell}\p{k_{p-1-r},q} \\
  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{\mu np\ell}\p{0,q} \\
  &= { \ell + p - 1 \choose p - 1 }
  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.
\end{align}
Substitution of this result together with $D_{\mu,m-p,n-p}^{(\ell)}$
with \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
\begin{align}
  E_{\mu mn}^{(p)}
  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{\mu mnp}
\end{align}
with
\begin{align}
  G_{\mu mnp}
  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
  \p{m-p-\ell} \p{\ell+m-p+F_{\mu np}}
  \p{2\ell + F_{\mu np}}
  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}} \\
  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{\mu np}} \\
  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-n-q+2\mu S_z} \\
  &= { m \choose p + 1 } \sum_{q=0}^{p+1} \p{2\mu S_z}^q
  e_{p+1-q}\p{m-n-\Z_{p+1}},
\end{align}
where $\Z_k$ denotes the set of $k$ smallest non-negative integers;
$m-n-\Z_{p+1}\equiv\set{m-n-j:j\in\Z_{p+1}}$; and $e_k\p{\S}$ is the
elementary symmetric polynomial of degree $k$ in elements of $\S$,
i.e.~the sum of all products of $k$ distinct elements chosen from
$\S$.  Among other ways, the elementary symmetric polynomial
$e_k\p{\S_n}$ for $\S_n\equiv\set{x_1,x_2,\cdots,x_n}$ can be computed
via the recurrence relation
\begin{align}
  e_k\p{\S_n} = e_k\p{\S_{n-1}} + x_n e_{k-1}\p{\S_{n-1}}
\end{align}
together with the terminal conditions $e_k\p{\S_k}=\prod_{j=1}^k x_j$
and $e_0\p{\S_n}=1$.  Finally, we can say that
\begin{align}
  E_{\mu mn}^{(p-1)}
  = p! { m \choose p } { n \choose p }
  S_\bmu^{n-p} S_\mu^{m-p}
  \sum_{q=0}^p \p{2\mu}^q e_{p-q}\p{m-n-\Z_p} S_z^q,
\end{align}
where in this form $E_{\mu mn}^{(0)} = D_{\mu mn}$, so
\begin{align}
  \sp{S_\mu^m, S_\bmu^n}_-
  = \sum_{p=1}^{\min\set{m,n}} E_{\mu mn}^{(p-1)}.
\end{align}

>>>>

where the first two terms in the sum over $p$ are

@@

  &= n \sum_{k=0}^{m-2} D_{\mu,m-1,n-1}^{(k)}
  \p{2k+1-n+2\mu S_z}, \\

@@

  &= n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}

@@

\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} can be used to evaluate the
commutator $\sp{S_\mu^m,S_\nu^n}_- = C_{\mu mn}$, but these result are
-- to put it lightly -- quite a mess: the expression for

@@

Observing that we always have
$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$ in \eqref{eq:E_mn_p}, we
can rearrange the order of the sums and relabel $k_p\to\ell$ to get
\begin{align}
  E_{\mu mn}^{(p)}
  = \f{n!}{\p{n-p}!}
  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+F_{\mu np}}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}},
  \label{eq:E_mn_p_sum}
\end{align}
where for shorthand we define
\begin{align}
  F_{\mu np} \equiv p - n + 2\mu S_z,
  &&
  \sum_{\p{\v k,p,\ell}} X \equiv
  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
  \cdots \sum_{k_{p-1}=k_{p-2}}^\ell X.
\end{align}
We now define
\begin{align}
  f_{\mu np\ell}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{\mu np}},
\end{align}
and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
finding
\begin{align}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  &= \sum_{\p{\v k,p-1,\ell}}
  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  f_{\mu np\ell}\p{k_{p-2},1} \\
  &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  \prod_{q=1}^r f_{\mu np\ell}\p{k_{p-1-r},q} \\
  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{\mu np\ell}\p{0,q} \\
  &= { \ell + p - 1 \choose p - 1 }
  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.
\end{align}
Substitution of this result together with $D_{\mu,m-p,n-p}^{(\ell)}$
with \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
\begin{align}
  E_{\mu mn}^{(p)}
  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{\mu mnp}
\end{align}
with
\begin{align}
  G_{\mu mnp}
  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
  \p{m-p-\ell} \p{\ell+m-p+F_{\mu np}}
  \p{2\ell + F_{\mu np}}
  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}} \\
  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{\mu np}}.
\end{align}
where we can expand
\begin{align}
  \prod_{q=0}^p \p{m-p-q+F_{\mu np}}
  = \sum_{q=0}^{p+1} \p{2\mu S_z}^q
  = \sum_{q=0}^{p+1} \p{2\mu S_z}^q
  e_{p+1-q}\p{m-n-\Z_{p+1}}.
\end{align}
Here $\Z_k$ denotes the set of $k$ smallest non-negative integers;
$m-n-\Z_{p+1}\equiv\set{m-n-j:j\in\Z_{p+1}}$; and $e_k\p{\S}$ is the
elementary symmetric polynomial of degree $k$ in elements of $\S$,
i.e.~the sum of all products of $k$ distinct elements chosen from
$\S$.  The elementary symmetric polynomial $e_k\p{\S_n}$ for
$\S_n\equiv\set{x_1,x_2,\cdots,x_n}$ can be computed (among other
ways) via the recurrence relation
\begin{align}
  e_k\p{\S_n} = e_k\p{\S_{n-1}} + x_n e_{k-1}\p{\S_{n-1}}
\end{align}
together with the terminal conditions $e_k\p{\S_k}=\prod_{j=1}^k x_j$
and $e_0\p{\S_n}=1$.  Finally, we can say that
\begin{align}
  E_{\mu mn}^{(p-1)}
  = p! { m \choose p } { n \choose p }
  S_\bmu^{n-p} S_\mu^{m-p}
  \sum_{q=0}^p \p{2\mu}^q e_{p-q}\p{m-n-\Z_p} S_z^q,
\end{align}
where in this form $E_{\mu mn}^{(0)} = D_{\mu mn}$, so
\begin{align}
  \sp{S_\mu^m, S_\bmu^n}_-
  = \sum_{p=1}^{\min\set{m,n}} E_{\mu mn}^{(p-1)}.
\end{align}
